---
title: "ps_4"
author: "Sabrina Chok"
date: "February 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #basic necessity  
library(readxl) #read excel files
library(ggplot2) #create graphs from data
library(devtools) #make nice looking table
library(gt)

```

```{r, echo = FALSE, include=FALSE}
#reads csv file and puts dataset into a variable
x <- read_csv("ps_4_elections-poll-nc09-3.csv") 
```

# Question 1

```{r, echo = FALSE}
#find # of respondents who supported the Democratic candidate.
ml1 <- x %>% 
  filter(response == "Dem") %>% #filters for responses marked as Democratic
  count() #counts number of responses
```

There were `r ml1` respondents who supported the Democratic candidate.

```{r, echo = FALSE}
#find difference between the # of respondents who favored the Republican candidate and those who were Undecided.
ml2 <- x %>% 
  count(response) %>% #counts number of responses for each party
  spread(response, n) %>% #spreads parties across columns for easy arithmetic in next step
  mutate(diffrep_und = Rep - Und) #create variable with difference between republican and undecided
```

There were `r ml2$diffrep_und` more respondents who favored the Republican candidate than who were Undecided.

```{r, echo = FALSE}
#find # of individuals with different values in the two gender variables (gender and gender_combined)
ml3 <- x %>% 
  select(gender, gender_combined) %>% #select columns with gender information
  filter(gender != gender_combined) %>% #filter out rows in which values in two columns are different
  count() #count the number of rows to find answer
```

There are two gender variables (gender and gender_combined). There are `r ml3` individuals for whom these variables have different values.

```{r, echo = FALSE}
#There are X respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.
ml4 <- x %>% 
  select(file_race_black, race_eth) %>% #select columns with race information
  filter(race_eth == "White") %>% #filter out respondents listed as "white" under race_eth
  filter(file_race_black != race_eth) %>% #filter out rows with with different values between two columns
  count() #count total number of rows to find answer

```

There are `r ml4` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

```{r, echo = FALSE}
#The first response of Dem came X minutes (rounded to the nearest minute) before the first response of Rep.
ml5 <- x %>% 
  select(response, timestamp) %>% #select columns with party and timestamp information
  filter(response %in% c("Dem", "Rep")) %>% #filter out democrat and republican responses
  group_by(response) %>% #group data by response 
  summarize(firstresp = min(timestamp)) #find earliest timestamp for each response

#find difference between first dem and rep response, takes out numeric part of output, rounds it to nearest whole number
a <- round(as.numeric(ml5[2,2] - ml5[1,2]), digit = 0)
```

The first response of Dem came `r a` minutes (rounded to the nearest minute) before the first response of Rep.

# Question 2
```{r, echo = FALSE}
table <- x %>% 
  select(response, race_eth, final_weight) %>% #select response, race, and response columns
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>%  #filter out unnecessary data
  mutate(race_eth = fct_relevel(race_eth, "White", "Black", "Hispanic", "Asian", "Other")) %>% #reorder ethnicities to match the desired table
  group_by(race_eth, response) %>% #group by race and response
  summarize(total = sum(final_weight)) %>% #create variable, total, which sums up groups based on final weights
  spread(key = response, value = total, fill = 0) %>% #spread data by response and total number of respondents
  mutate(all = Dem + Rep + Und +`3`) %>% #create new column that sums total number of respondents
  mutate(Dem = Dem / all) %>% #divides individual values on table by total to find ratios
  mutate(Rep = Rep / all) %>% 
  mutate(Und = Und / all) %>% 
  select(-all, -`3`) %>% #hide columns with data not necessary for final table
  ungroup() #ungroup data 

gt(table) %>% 
  #add title and subtitle
  tab_header(
    title = "2018 North Carolina's 9th Congressional District Polling Data", 
    subtitle = "New York Times Upshot/Siena College 2018 live polls") %>% 
  #rename columns
  cols_label(
      race_eth = "",
      Dem = "DEM.",
      Rep = "REP.",
      Und = "UND.") %>% 
  #change numbers in table to percents
  fmt_percent(columns = vars(Dem, Rep, Und), decimal = 0) %>% 
  #turns missing values into dashes
  na_if(0) %>% 
  fmt_missing(columns = 4)
```


# Question 3
```{r, echo = FALSE}
x %>% 
  select(educ, final_weight) %>% #select education and final weight data
  filter(educ != "[DO NOT READ] Refused") %>% #filter out irrelevant data
  mutate(educ = fct_relevel(educ, "Grade school", "High school", "Some college or trade school", 
                            "Bachelors' degree", "Graduate or Professional Degree")) %>% #reorder education to match the desired table
  #create violin plot with jitters to match the actual graph; varied alpha, width, and size until desired aesthetic appearance
  ggplot(aes(x = educ, y = final_weight)) + geom_violin() + geom_jitter(alpha = .5, width = .2, size = .9) + coord_flip() +
  #add title, subtitle, caption, and graph axis labels
  labs(title = "More Educated Matter Less in North Carolina 9th",
       subtitle = "Poll gives more weight to people who are less likely to participate in polls", 
       caption = "New York Times Upshot/Siena College 2018 live polls") +
       xlab(NULL) +
       ylab("Weight Given to Respondent in Calculating Poll Results")
```

# Question 4
```{r, echo = FALSE}

#create facet of bar graphs that compare respondent gender differences across parties in NC 9th
x %>% 
  filter(partyid != "[DO NOT READ] Refused", partyid != "or as a member of another political party") %>% #filter out undesired data to focus only on dems, rep, and independents
  select(gender, partyid) %>% #select gender and party identification of respondents
  group_by(partyid, gender) %>% #group by party identification and gender
  count() %>% #count number of respondents for each set of party and gender
  #create bar plot with gender on x axis, faceted by party identification
  ggplot(aes(x = gender, y = n, fill = gender)) + geom_col() + facet_grid(~partyid)+ 
  #labeled plot with appropriate title and caption; left out legend and axis labels for aesthetic purposes (but also they weren't necessary to understand the graph)
  labs(title = "Party Identification by Gender in North Carolina 9th",
       caption = "New York Times Upshot/Siena College 2018 live polls") +
  xlab(NULL) +
  ylab(NULL) +
  theme(legend.position = 'none')
```

#Colleagues
Jenny Li, Henry Zhu, Neil Khurana, Albert Shin, Mike Silva 